---
title: Homework 4
subtitle: Data Visualization and Data Transformation
date: last-modified
from: markdown+emoji
comments: false
code-fold: false

execute: 
  warning: false
  message: false
  fig-width: 9
  fig-height: 7
  fig-align: center
---
```{r}
#| eval: true
#| include: false

library(tidyverse)
library(skimr)
```

# Direction

- Download the Homework 4 Quarto document, `danl-200-hw-04-q.qmd`, by clicking [this link](https://drive.google.com/file/d/1NT5nKWEjwi536RYxIkoZrdUVfDzllA54/view?usp=share_link).

- Provide both your answer to each sub question in **Question 1 of Homework 4** in the `danl-200-hw-04-q.qmd`.

- Please submit your **Quarto document for Question 1 of Homework 4** to the Brightspace with the name below:

  - `danl-200-hw4-LASTNAME-FIRSTNAME.qmd`\
  ( e.g., `danl-200-hw4-choe-byeonghak.qmd` )

- The due is April 30, 2024, 11:59 P.M.

- Please send an email to Byeong-Hak (`bchoe@geneseo.edu`) if you have any questions or clarifications.

<br><br><br>

## Question 1

Load the data.frame for Question 1.

```{r}
#| echo: true
#| eval: true
restaurant <- read_csv('https://bcdanl.github.io/data/DOHMH_NYC_Restaurant_Inspection.csv')
```

```{r}
#| results: asis
#| echo: false
rmarkdown::paged_table(restaurant)
```

### Variable Description

-   `CAMIS`:
    -   This is an unique identifier for the entity (restaurant);
    -   10-digit integer
-   `DBA`:
    -   This field represents the name (doing business as) of the entity (restaurant);
    -   Public business name, may change at discretion of restaurant owner
-   `BORO`:
    -   Borough in which the entity (restaurant) is located.;
    -   • `1` = MANHATTAN
    -   • `2` = BRONX
    -   • `3` = BROOKLYN
    -   • `4` = QUEENS
    -   • `5` = STATEN ISLAND
    -   • `0` = Missing;
-   `CUISINE DESCRIPTION`:
    -   This field describes the entity (restaurant) cuisine.
-   `ACTION`:
    -   This field represents the actions that is associated with each restaurant inspection. ;
    -   • Violations were cited in the following area(s).
    -   • No violations were recorded at the time of this inspection.
    -   • Establishment re-opened by DOHMH
    -   • Establishment re-closed by DOHMH
    -   • Establishment Closed by DOHMH.
    -   • Violations were cited in the following area(s) and those requiring immediate action were addressed.
-   `VIOLATION CODE`:
    -   Violation code associated with an establishment (restaurant) inspection
-   `VIOLATION DESCRIPTION`:
    -   Violation description associated with an establishment (restaurant) inspection
-   `CRITICAL FLAG`:
    -   Indicator of critical violation;
    -   • `Critical`
    -   • `Not Critical`
    -   • `Not Applicable`;
    -   Critical violations are those most likely to contribute to food-borne illness
-   `SCORE`:
    -   Total score for a particular inspection;
-   `GRADE`:
    -   Grade associated with the inspection;
    -   • `N` = Not Yet Graded
    -   • `A` = Grade A
    -   • `B` = Grade B
    -   • `C` = Grade C
    -   • `Z` = Grade Pending
    -   • `P` = Grade Pending issued on re-opening following an initial inspection that resulted in a closure

<br><br>

### Q1a.

What are the mean, standard deviation, first quartile, median, third quartile, and maximum of `SCORE` for each `GRADE` of restaurants?

*Answer*:
```{r}
skim(restaurant)
```

<br><br>

### Q1b.

- How many restaurants with a `GRADE` of `A` are there in each `BORO` in NYC?

*Answer*:
```{r}
Q1b <- restaurant |> 
  select(GRADE, 
         BORO) |> 
  arrange(GRADE)
```

<br>

### Q1c.

- Provide both (1) ggplot code and (2) a simple comment to describe how the distribution of `SCORE` varies by `GRADE` and `CRITICAL FLAG`.



*Answer*:
```{r}
Q1c <- restaurant |> 
  select(GRADE,
         SCORE,
         `CRITICAL FLAG`)

ggplot(Q1c,
       aes(x = GRADE, 
           y = 'CRITICAL FLAG')) +
  geom_point() +
  facet_wrap(SCORE ~ .)

#The score has higher critical flags for higher grades the higher that it goes, at 0 all A, B, and C grades have critical flags
```

<br><br>

### Q1d.

Provide both (1) ggplot code and (2) a simple comment to describe how the proportion of `CRITICAL FLAG` varies by `GRADE` and `BORO`.


*Answer*:
```{r}
Q1d <- restaurant |> 
  select('CRITICAL FLAG',
         GRADE,
         BORO)

ggplot(Q1d,
       aes(x = GRADE, 
           y = 'CRITICAL FLAG'),
       alpha = .01) +
  geom_point() +
  facet_wrap(BORO ~ .,
             scales = "free_x")

# There is a heavy amount of critical flags per Boro for all the grades
```

<br><br>

### Q1e.
- Calculate the proportion of restaurants that had violation related to `contamination` or `mice` in each `BORO`.

*Answer*:
```{r}
Q1e <- restaurant |> 
  summarise('MICE', 'CONTAMINATION', BORO)
```

<br><br>


### Q1f.

-   Find the 3 most common names of restaurants (`DBA`) in each `BORO`.
-   Overall, which `DBA` value is most common in NYC?


*Answer*:
```{r}
Q1f <- restaurant |> 
  select(DBA, BORO) |> 
  group_by(DBA, BORO) |> 
  summarise(n = n()) |> 
  arrange(-n)

#The three most common restaurant names are Dunkin, Starbucks, and McDonalds
```

<br><br>


### Q1g.

- Construct a bar chart to display the top 10 names of restaurants (`DBA`) sortetd by the frequency of each name within their respective borough (`BORO`).


*Answer*:
```{r}
ggplot(Q1f,
       aes(x = DBA))+
  geom_bar() +
  facet_wrap(BORO ~ .,
             scales = "free_y")
```

<br><br>


### Q1h
What is your GitHub username?


*Answer*:
```{r}
ccough77
```


<br><br><br>

## Question 2

- Create a blog post featuring data analysis on the `health_cust` data.frame and publish it on your GitHub website using your Quarto document for Homework 2.
  - If you have not completed on Homework 2, please take this opportunity to do so.

```{r}
health_cust <- read_csv(
  'https://bcdanl.github.io/data/custdata_rev.csv'
)
```


<br><br>

